---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# fwildclusterboot

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/fwildclusterboot)](https://CRAN.R-project.org/package=fwildclusterboot)
<!-- badges: end -->

The goal of fwildclusterboot is to estimate a fast wild cluster bootstrap for linear regression models of classes "lm", "lm_robust" and "felm" (from packages base, estimatr and lfe).

## Installation

You can install the released version of fwildclusterboot from github...

``` r
install.packages("fwildclusterboot")
```

```{r, echo=FALSE, include=FALSE, results="hide"}
# install the R package
#load_all("fwildclusterboot")
# library("devtools")
# setwd("C:/Users/au563468/Dropbox")
# install("fwildclusterboot")
# document("fwildclusterboot")

```
```{r, echo=FALSE, include=FALSE, results="hide"}
# bothPaths <- .libPaths()   # extract both paths
# bothPaths <- c(bothPaths [2], bothPaths [1])  # change order
# .libPaths(bothPaths )  # modify the order
# setwd("C:/Users/au563468/Dropbox")
# devtools::install("fwildclusterboot")
# devtools::document("fwildclusterboot")
# 
# library(usethis)
# setwd("C:/Users/au563468/Dropbox/fwildclusterboot")
# use_package("data.table", type = "Imports")
# use_package("estimatr", type = "Suggests")
# use_package("magrittr", type = "Suggests")
# use_package("mvtnorm", type = "Suggests")
# use_package("lmtest", type = "Suggests")
# use_package("lfe", type = "Suggests")

library(data.table)
library(estimatr)
library(magrittr)
library(mvtnorm)
library(multiwayvcov)
library(lmtest)
library(lfe)
library(pracma)
library(fabricatr)
library(rbenchmark)
library(fixest)
```

In a first step, simulate some data

```{r}

library(fwildclusterboot)
B <- 10000
seed <- 1345671
set.seed(seed)


voters <- fabricate(
  N = 2000,
  group_id = rep(1:20, 100),
  ideology = draw_normal_icc(mean = 0, N = N, clusters = group_id, ICC = 0.1),
  ideological_label = draw_ordered(
    x = ideology,
    break_labels = c(
      "Very Conservative", "Conservative",
      "Liberal", "Very Liberal"
    )
  ),
  income = exp(rlnorm(n = N, meanlog = 2.4 - (ideology * 0.1), sdlog = 0.12)),
  Q1_immigration = draw_likert(x = ideology, type = 7),
  Q2_defence = draw_likert(x = ideology + 0.5, type = 7),
  treatment = draw_binary(0.5, N = N),
  proposition_vote = draw_binary(latent = ideology + 0.01 * treatment, link = "probit")
)

setDT(voters)
voters[, log_income := log(income)]
voters[, Q1_immigration := as.factor(Q1_immigration) ]
voters[, Q2_defence := as.factor(Q2_defence)]

head(voters)

```

The fwildclusterboot package supports estimation of linear models based on base R's lm() function, estimatr's lm_robust function, 
lfe's felm() function and fixest's feols() function. 

```{r}
library(estimatr)
library(lfe)
library(fixest)

lm_fit <- lm(proposition_vote ~ treatment + ideology + log_income +Q1_immigration + Q2_defence, weights = NULL, data = voters)
lm_robust_fit <- lm_robust(proposition_vote ~ treatment + ideology + log_income, fixed_effects = ~ Q1_immigration + Q2_defence, weights = NULL, data = voters)
lm_robust_fit1 <- lm_robust(proposition_vote ~ treatment + ideology + log_income + Q1_immigration + Q2_defence, weights = NULL, data = voters )
feols_fit <- feols(proposition_vote ~ treatment + ideology + log_income, fixef = c("Q1_immigration", "Q2_defence"), weights = NULL, data = voters)
felm_fit <- felm(proposition_vote ~ treatment + ideology + log_income | Q1_immigration + Q2_defence, weights = NULL, data = voters)
```

The boottest command offers two functions. First, it calculates p-values for a given null hypothesis. This is ususally extremely fast: 

```{r}
lm = boottest.lm(lm_fit, clustid = voters$group_id, B = B, seed = seed, param = "treatment", conf_int = FALSE)
estimatr_fe = boottest.lm_robust(lm_robust_fit, clustid = voters$group_id, B = B, seed = seed, param = "treatment", conf_int = FALSE)
estimatr = boottest.lm_robust(lm_robust_fit1, clustid = voters$group_id, B = B, seed = seed, param = "treatment", conf_int = FALSE)
felm = boottest.felm(felm_fit, clustid = voters$group_id, B = B, seed = seed, param = "treatment", conf_int = FALSE)
fixest = boottest.fixest(feols_fit, clustid = voters$group_id, B = B, seed = seed, param = "treatment", conf_int = FALSE)
```

```{r, echo=FALSE, include=FALSE, results="hide"}
# debugging: 
# object <- lm_fit
# clustid = voters$group_id
# B = B
# seed = seed
# param = "treatment"
# conf_int = FALSE

```


Secondly, the user may specify to obtain confidence intervals. 

```{r}
res_lm = boottest.lm(lm_fit, clustid = voters$group_id, B = B, seed = seed, param = "treatment", conf_int = TRUE)
#res_estimatr_fe = boottest.lm_robust(lm_robust_fit, clustid = voters$group_id, B = B, seed = seed, param = "treatment", conf_int = #TRUE)
#res_estimatr = boottest.lm_robust(lm_robust_fit1, clustid = voters$group_id, B = B, seed = seed, param = "treatment", conf_int = #TRUE)
res_felm = boottest.felm(felm_fit, clustid = voters$group_id, B = B, seed = seed, param = "treatment", conf_int = TRUE)
res_fixest = boottest.fixest(feols_fit, clustid = voters$group_id, B = B, seed = seed, param = "treatment", conf_int = TRUE)
```

This leads to the following results: 

```{r}
summary(res_lm)
#summary(res_estimatr_fe)
#summary(res_estimatr)
summary(res_felm)
summary(res_fixest)
```





