---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# fwildclusterboot

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/fwildclusterboot)](https://CRAN.R-project.org/package=fwildclusterboot)
<!-- badges: end -->

The goal of fwildclusterboot is to estimate a fast wild cluster bootstrap for linear regression models of classes "lm", "lm_robust" and "felm" (from packages base, estimatr and lfe).

## Installation

You can install the released version of fwildclusterboot from github...



```{r, echo=FALSE, include=FALSE, results="hide"}
# bothPaths <- .libPaths()   # extract both paths
# bothPaths <- c(bothPaths [2], bothPaths [1])  # change order
# .libPaths(bothPaths )  # modify the order
#  setwd("C:/Users/au563468/Dropbox")
#  devtools::install("fwildclusterboot")
#  devtools::document("fwildclusterboot")
# # 
#  library(usethis)
#  setwd("C:/Users/au563468/Dropbox/fwildclusterboot")
#  use_package("data.table", type = "Imports")
#  use_package("estimatr", type = "Imports")
# 
#  use_package("fixest", type = "Imports")
#  use_package("magrittr", type = "Suggests")
#  use_package("fabricatr", type = "Suggests")
#  use_package("lmtest", type = "Suggests")
#  use_package("lfe", type = "Suggests")


library(data.table)
library(multiwayvcov)
library(lmtest)
library(pracma)
library(fabricatr)
library(rbenchmark)
library(fixest)
 
setwd("C:/Users/au563468/Dropbox/fwildclusterboot/R")
file.sources = list.files(pattern="*.R")
sapply(file.sources, source, .GlobalEnv)

```

In a first step, simulate some data

```{r}


# library(fwildclusterboot)
B <- 10000
seed <- 1345671
set.seed(seed)

voters <- create_data(N = 2000, N_G = 20)
head(voters)

```

The fwildclusterboot package supports estimation of linear models based on base R's lm() function, estimatr's lm_robust function, 
lfe's felm() function and fixest's feols() function. 

```{r}
library(estimatr)
library(lfe)
library(fixest)

lm_fit <- lm(proposition_vote ~ treatment + ideology + log_income +Q1_immigration + Q2_defence, weights = NULL, data = voters)
lm_robust_fit <- lm_robust(proposition_vote ~ treatment + ideology + log_income, fixed_effects = ~ Q1_immigration + Q2_defence, weights = NULL, data = voters)
lm_robust_fit1 <- lm_robust(proposition_vote ~ treatment + ideology + log_income + Q1_immigration + Q2_defence, weights = NULL, data = voters )
feols_fit <- feols(proposition_vote ~ treatment + ideology + log_income, fixef = c("Q1_immigration", "Q2_defence"), weights = NULL, data = voters)
felm_fit <- felm(proposition_vote ~ treatment + ideology + log_income | Q1_immigration + Q2_defence, weights = NULL, data = voters)
```

The boottest command offers two functions. First, it calculates p-values for a given null hypothesis of the form HO: $\beta = 0$ vs H1: $\beta \neq 1$. 
In order to work, an object from a regression model of class lm, lm_robust, felm or feols needs to be passed to the boottest function. Currently, the user is still required to pass a vector with information on clusters to the function. As of now, the function only supports one-dimensional clustering. 
By default, the boottest function calculates confidence intervals by inversion. The user can considerably speed up the inference procedure by setting the argument conf_int to FALSE, in which case no confidence intervals are computed.

```{r}
lm = boottest(lm_fit, clustid = voters$group_id, B = B, seed = seed, param = "treatment", conf_int = FALSE)
estimatr_fe = boottest(lm_robust_fit, clustid = voters$group_id, B = B, seed = seed, param = "treatment", conf_int = FALSE)
estimatr = boottest(lm_robust_fit1, clustid = voters$group_id, B = B, seed = seed, param = "treatment", conf_int = FALSE)
felm = boottest(felm_fit, clustid = voters$group_id, B = B, seed = seed, param = "treatment", conf_int = FALSE)
fixest = boottest(feols_fit, clustid = voters$group_id, B = B, seed = seed, param = "treatment", conf_int = FALSE)
```

```{r, echo=FALSE, include=FALSE, results="hide"}
# debugging: 
# object <- lm_fit
# clustid = voters$group_id
# B = B
# seed = seed
# param = "treatment"
# conf_int = FALSE

```


Secondly, the user may specify to obtain confidence intervals. 

```{r}
res_lm = boottest(lm_fit, clustid = voters$group_id, B = B, seed = seed, param = "treatment", conf_int = TRUE)
#res_estimatr_fe = boottest(lm_robust_fit, clustid = voters$group_id, B = B, seed = seed, param = "treatment", conf_int = #TRUE)
#res_estimatr = boottest(lm_robust_fit1, clustid = voters$group_id, B = B, seed = seed, param = "treatment", conf_int = #TRUE)
res_felm = boottest(felm_fit, clustid = voters$group_id, B = B, seed = seed, param = "treatment", conf_int = TRUE)
res_fixest = boottest(feols_fit, clustid = voters$group_id, B = B, seed = seed, param = "treatment", conf_int = TRUE)
```

A summary method collects the results. 

```{r}
summary(res_lm)
#summary(res_estimatr_fe)
#summary(res_estimatr)
summary(res_felm)
summary(res_fixest)
```

Furthermore, boottest comes with a tidy method which, in analogy with the broom-package, returns the estimation results as a data.frame.

```{r}
tidy(res_lm)
#summary(res_estimatr_fe)
#summary(res_estimatr)
tidy(res_felm)
tidy(res_fixest)
```

## Benchmarks

```{r}

# seed <- 1
# set.seed(seed)
# N <- 5000
# N_G <- 10
# B <- 1000
# 
# data <- create_data(N = N, N_G = N_G)
# 
# lapply(c(100, 1000, 5000, 1000), function(B){
#       lm_fit <- lm(proposition_vote ~ treatment + ideology + log_income +Q1_immigration + Q2_defence, weights = NULL, data = data)
#       bench <- benchmark(
#           boot =  multiwayvcov::cluster.boot(lm_fit, 
#                                              as.factor(data$group_id), 
#                                              R = B, 
#                                              boot_type = "residual", 
#                                              wild_type = "rademacher", 
#                                              parallel = FALSE), 
#           fast_boot_1 = boottest(lm_fit, clustid = data$group_id, B = B, seed = seed, param = "treatment", conf_int = FALSE),
#           fast_boot_1_2 = boottest(lm_fit, clustid = data$group_id, B = B, seed = seed, param = "treatment", conf_int = TRUE), 
#           replications = 10)
#       c(N, N_G, bench$elapsed)
# })
# 
# 
# res <- data.frame()
# 
# data <- create_data(N = N, N_G = 20)
# i <- 0
# 
# res <- 
# lapply(c(1000, 5000, 10000, 20000), function(N){
#   lapply(c(10, 20, 50, 100, 200), function(N_G){
#     data <- create_data(N = N, N_G = N_G)
#     lapply(c(100, 1000, 5000, 1000), function(B){
#       lm_fit <- lm(proposition_vote ~ treatment + ideology + log_income +Q1_immigration + Q2_defence, weights = NULL, data = data)
#       bench <- benchmark(
#         boot =  multiwayvcov::cluster.boot(lm_fit, 
#                                                        as.factor(data$group_id), 
#                                                        R = B, 
#                                                        boot_type = "residual", 
#                                                        wild_type = "rademacher", 
#                                                        parallel = FALSE), 
#         fast_boot_1 = boottest(lm_fit, clustid = data$group_id, B = B, seed = seed, param = "treatment", conf_int = FALSE),
#         fast_boot_1_2 = boottest(lm_fit, clustid = data$group_id, B = B, seed = seed, param = "treatment", conf_int = TRUE), 
#         replications = 10, 
#         columns = c("bootstrap", "fast bootstrap 1", "fast bootstrap 2")#,
#         #relative = c("bootstrap")
#       )
#       c(N, N_G, bench$elapsed)
#     })
#   })
# })





```






