---
output: github_document
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# fwildclusterboot


<!-- badges: start -->
<!-- [![packageversion](https://img.shields.io/badge/Package%20version-`r version`-orange.svg?style=flat-square)](commits/master) -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/fwildclusterboot)](https://CRAN.R-project.org/package=fwildclusterboot)
[![R-CMD-check](https://github.com/s3alfisc/fwildclusterboot/workflows/R-CMD-check/badge.svg)](https://github.com/s3alfisc/fwildclusterboot/actions)
[![Codecov test coverage](https://codecov.io/gh/s3alfisc/fwildclusterboot/branch/master/graph/badge.svg)](https://codecov.io/gh/s3alfisc/fwildclusterboot?branch=master)
[![minimal R version](https://img.shields.io/badge/R%3E%3D-4.0.0-6666ff.svg)](https://cran.r-project.org/)

<!-- badges: end -->

The `fwildclusterboot` package is an R port of Stata's [boottest](https://github.com/droodman/boottest) package. 

It implements the fast wild cluster bootstrap algorithm developed in [Roodman et al (2019)](https://econpapers.repec.org/paper/qedwpaper/1406.htm) for regression objects in R. It currently works for regression objects of type `lm`, `felm` and `fixest` from base R and the `lfe` and `fixest` packages. 

The package's central function is `boottest()`. It allows the user to test two-sided, univariate hypotheses using a wild cluster bootstrap. Importantly, it uses the "fast" algorithm developed in Roodman et al, which makes it feasible to calculate test statistics based on a large number of bootstrap draws even for large samples -- as long as the number of bootstrapping clusters is not too large.

The `fwildclusterboot` package currently supports multi-dimensional clustering and one-dimensional, two-sided hypotheses. It supports regression weights, multiple distributions of bootstrap weights, fixed effects, restricted (WCR) and unrestricted (WCU) bootstrap inference and subcluster bootstrapping for few treated clusters [(MacKinnon & Webb, (2018))](https://academic.oup.com/ectj/article-abstract/21/2/114/5078969).

<!-- The following features will be added in the future:  -->

<!-- * support for multivariate hypotheses  -->
<!-- * bootstrap distributions beyond the rademacher distribution -->


```{r, echo=FALSE, include=FALSE, results="hide", execute = FALSE}
# setwd("C:/Users/alexa/Dropbox")
# devtools::install("fwildclusterboot")
# devtools::document("fwildclusterboot")
# devtools::test("C:/Users/alexa/Dropbox/fwildclusterboot")
# usethis::use_github_actions()
# usethis::use_github_action(“render-readme.yaml”)
#usethis::use_github_actions_badge(name = "R-CMD-check", repo_spec = NULL)#

# usethis::use_mit_license("C:/Users/alexa/Dropbox/fwildclusterboot")
# # # cmd check r 
# # Sys.setenv('_R_CHECK_SYSTEM_CLOCK_' = 0)
# # #system("defaults write org.R-project.R force.LANG en_US.UTF-8") # restart R
# # Sys.getlocale()
# # Sys.setlocale("LC_MESSAGES", "C")
# # #Sys.setenv(LANG = "en_US.UTF-8")
# Sys.getlocale()

# # install
# setwd("C:/Users/alexa/Dropbox")
# devtools::install("fwildclusterboot")
# devtools::document("fwildclusterboot")
# devtools::test("C:/Users/alexa/Dropbox/fwildclusterboot")
# devtools::test("C:/Users/alexa/Dropbox/fwildclusterboot")
# devtools::check("C:/Users/alexa/Dropbox/fwildclusterboot")



# execute all functions in fwildclusterboot 
# setwd("C:/Users/alexa/Dropbox/fwildclusterboot/R")
# file.sources = list.files(pattern="*.R")
# sapply(file.sources, source, .GlobalEnv)

```


## Benchmarks

Results of timing benchmarks of `fwildclusterboot` with `sandwich::vcovBS` (one replication):

* Benchmark 1: one cluster with dimension $N_G = 40$
* Benchmark 2: two closters with dimensions $N_G1= 40$, $N_G2 = 20$, $N_G12 = 800$

```{r,echo=FALSE, warning = FALSE, message = FALSE}

benchmarks <- readRDS("C:/Users/alexa/Dropbox/fwildclusterboot develop/benchmarks_new.rds")
library(data.table)

setDT(benchmarks)

benchmarks[benchmarks$test == "fwc_1", test := "boottest, no fixed effect"]
benchmarks[benchmarks$test == "fwc_2", test := "boottest, fixed effect"]
benchmarks[benchmarks$test == "sw_1c",test :="sandwich, 1 core"]
benchmarks[benchmarks$test == "sw_2c",test :="sandwich, 2 cores"]
benchmarks[benchmarks$test == "sw_4c",test := "sandwich, 4 cores"]


library(ggplot2)
library(scales)
plot <- 
ggplot(aes(x = B, y = elapsed, group = test, color = test), data = benchmarks) + 
  facet_wrap(~ test_nr, nrow = 2) + 
  scale_y_continuous(trans = 'log10', labels = comma) +
  geom_point() + 
  geom_line(alpha = 0.5) + 
  xlab("Number of Bootstrap Iterations") + 
  ylab("Time in seconds, log scale") + 
  theme_bw() +
  theme(legend.title=element_blank())
  #theme(legend.position="bottom")  
plot

```

<!-- ## The `boottest` function  -->


<!-- The `fwildclusterboot` package supports wild cluster bootstrap inference for linear models based on  -->

<!-- - `lm()` from `base` R -->
<!-- - `felm()` from `lfe` -->
<!-- - `feols()` from `fixest` -->

<!-- ```{r, warning = FALSE, message = FALSE} -->
<!-- library(fwildclusterboot) -->

<!-- B <- 99999 -->
<!-- seed <- 942413 -->
<!-- set.seed(seed) -->
<!-- voters <- create_data_2(N = 10000, N_G1 = 20, icc1 = 0.01, N_G2 = 10, icc2 = 0.01, numb_fe1 = 10, numb_fe2 = 10, seed = seed) -->

<!-- # estimate the regression model -->
<!-- lm_fit <- lm(proposition_vote ~ treatment + ideology1 + log_income + Q1_immigration , data = voters) -->

<!-- # bootstrap estimation -->
<!-- boot_lm <- boottest(lm_fit, clustid = "group_id1", B = B, seed = seed, param = "treatment", conf_int = TRUE) -->

<!-- # summarize the results -->
<!-- summary(boot_lm) -->
<!-- tidy(boot_lm) -->

<!-- ``` -->



## Installation

You can install the development version of `fwildclusterboot` from github by following the steps below.  

```{r, eval = FALSE}
# note: installation requires Rtools
library(devtools)
install_github("s3alfisc/fwildclusterboot")
```



