---
output: github_document
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# fwildclusterboot

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/fwildclusterboot)](https://CRAN.R-project.org/package=fwildclusterboot)
<!-- badges: end -->

The `fwildclusterboot` package is an R port of Stata's `boottest` package. 

It implements the fast wild cluster bootstrap algorithm developed in Roodman et al (2019) for regression objects in R. It currently works for regression objects of type `lm`, `felm` and `fixest` from base R and the `lfe` and `fixest` packages. 

The package's central function is `boottest()`. It allows the user to test two-sided, univariate hypotheses using a wild cluster bootstrap. Importantly, it uses the "fast" algorithm developed in Roodman et al, which makes it feasible to calculate test statistics based on a large number of bootstrap draws even for large samples--as long as the number of bootstrapping clusters is not too large.

The `fwildclusterboot` package currently supports one- and two-dimensional clustering and one-dimensional hypotheses.

The package is highly experimental and only includes few unit tests.

The following features will be added in the future: 

* support for multivariate hypotheses 
* bootstrap distributions beyond the rademacher distribution


```{r, echo=FALSE, include=FALSE, results="hide", execute = FALSE}
# setwd("C:/Users/alexa/Dropbox")
# devtools::install("fwildclusterboot")
# devtools::document("fwildclusterboot")
# devtools::test("C:/Users/alexa/Dropbox/fwildclusterboot")
# usethis::use_github_actions()
# usethis::use_github_action(“render-readme.yaml”)
#usethis::use_github_actions_badge(name = "R-CMD-check", repo_spec = NULL)#

# usethis::use_mit_license("C:/Users/alexa/Dropbox/fwildclusterboot")
# # # cmd check r 
# # Sys.setenv('_R_CHECK_SYSTEM_CLOCK_' = 0)
# # #system("defaults write org.R-project.R force.LANG en_US.UTF-8") # restart R
# # Sys.getlocale()
# # Sys.setlocale("LC_MESSAGES", "C")
# # #Sys.setenv(LANG = "en_US.UTF-8")
# Sys.getlocale()

# # install
# setwd("C:/Users/alexa/Dropbox")
# devtools::install("fwildclusterboot")
# devtools::document("fwildclusterboot")
# devtools::test("C:/Users/alexa/Dropbox/fwildclusterboot")
# devtools::test("C:/Users/alexa/Dropbox/fwildclusterboot")
# devtools::check("C:/Users/alexa/Dropbox/fwildclusterboot")



# execute all functions in fwildclusterboot 
# setwd("C:/Users/alexa/Dropbox/fwildclusterboot/R")
# file.sources = list.files(pattern="*.R")
# sapply(file.sources, source, .GlobalEnv)

```

## The `boottest()` function 


The `fwildclusterboot` package supports wild cluster bootstrap inference for linear models based on 

- `lm()` from `base` R
- `felm()` from `lfe`
- `feols()` from `fixest`

```{r, warning = FALSE, message = FALSE}
library(fwildclusterboot)

B <- 10000
seed <- 942413
set.seed(seed)
voters <- create_data_2(N = 10000, N_G1 = 20, icc1 = 0.01, N_G2 = 10, icc2 = 0.01, numb_fe1 = 10, numb_fe2 = 10, seed = seed)

lm_fit <- lm(proposition_vote ~ treatment + ideology1 + log_income + Q1_immigration , weights = NULL, data = voters)
```

The `boottest` function always calculates p-values for a given univariate hypothesis test.
Second and by default, the boottest function calculates confidence intervals by inversion of the p-value. The user can considerably speed up the inference procedure by setting the argument `conf_int = FALSE`, in which case no confidence intervals are computed.

```{r, warning = FALSE, message=FALSE}
# 1) boottest based on object of class lm
boot_lm = boottest(lm_fit, clustid = "group_id1", B = B, seed = seed, param = "treatment", conf_int = TRUE)
```

The function `summary` collects the results. 
Boottest further comes with a `tidy` method, which, in analogy with the `broom` package, returns the estimation results as a data.frame.

```{r, warning = FALSE, message=FALSE}
summary(boot_lm)
tidy(boot_lm)
```


```{r, include = FALSE}
boot_lm_5 = boottest(lm_fit, clustid = "group_id1", B = B, seed = seed, param = "treatment", conf_int = TRUE, beta = 0, alpha = 0.05)

boot_lm_20 = boottest(lm_fit, clustid = "group_id1", B = B, seed = seed, param = "treatment", conf_int = TRUE, beta = 0, alpha = 0.20)

summary(boot_lm_5)
summary(boot_lm_20)

#confint(feols_fit, "treatment", level = 0.95, se = "cluster", cluster = "group_id1")
#confint(feols_fit, "treatment", level = 0.80, se = "cluster", cluster = "group_id1")

```

```{r, out.width="50%", include = FALSE}
plot(boot_lm)
```


## Installation

You can install the released version of `fwildclusterboot` from github by running 

```{r, eval = FALSE}
devtools::install_github("s3alfisc/fwildclusterboot")
```


```{r, warning = FALSE, message=FALSE, include = FALSE}

## Comparison to `vcovBS()` from `sandwich`

#The `multiwayvcov` package offers an alternative implementation of the wild bootstrap. 
#As can be seen, `sandwich::vcovBS`, `sandwich::vcov` and `boottest()` produce similar #results: 
  
library(sandwich)
library(lmtest)

res <- vcovBS(lm_fit, cluster = ~ group_id1, parallel = FALSE, R = 1000, wild_type = "rademacher")

# 1) asymptotic sandwich standard errors 
vocv <- vcovCL(lm_fit, ~ group_id1)
head(coeftest(lm_fit, vcov))

# 1) results from sandwich's cluster bootstrap 
head(coeftest(lm_fit, res))

# 2) results from fwildclusterboot
summary(boot_lm)



```



```{r, include = FALSE}
## Some Tests with 2-way clustering 

library(sandwich)
library(lmtest)
rm(boot_lm); rm(boot_fixest); rm(boot_felm); rm(res)


boot_lm <-  boottest(lm_fit, clustid = c("group_id1", "group_id2"), B = 1000, seed = seed, param = "treatment", conf_int = TRUE)

summary(boot_lm)

vocv <- vcovCL(lm_fit, ~ group_id1 + group_id2)
coeftest(lm_fit, vcov)

vcovBS <- vcovBS(lm_fit, cluster = ~ group_id1, parallel = FALSE, R = 1000, wild_type = "rademacher")

head(coeftest(lm_fit, vcovBS))

```



## Benchmarks

Results of timing benchmarks of `fwildclusterboot` with `sandwich::vcovBS`.

* Benchmark 1: N = 10000, one cluster with $N_G = 40$
* Benchmark 2: N = 10000, two closters with $N_G1= 40$ and $N_G2 = 20$, $N_G12 = 800$

```{r,echo=FALSE, warning = FALSE, message = FALSE}

benchmarks <- readRDS("C:/Users/alexa/Dropbox/fwildclusterboot develop/benchmarks.rds")

benchmarks[benchmarks$test == "fwc_1"]$test <- "boottest without fixed effect"
benchmarks[benchmarks$test == "fwc_2"]$test <- "boottest with fixed effect"
benchmarks[benchmarks$test == "sw_1c"]$test <- "sandwich, 1 core"
benchmarks[benchmarks$test == "sw_2c"]$test <- "sandwich, 2 cores"
benchmarks[benchmarks$test == "sw_2c"]$test <- "sandwich, 4 cores"


library(ggplot2)
library(scales)
plot <- 
ggplot(aes(x = B, y = elapsed, group = test, color = test), data = benchmarks) + 
  facet_wrap(~ test_nr) + 
 scale_y_continuous(trans = 'log10', labels = comma) + 
  geom_point() + 
  geom_line() + 
  xlab("Number of Bootstrap Iterations") + 
  ylab("Time in seconds, log scale") + 
  theme_bw()
plot

```

