---
title: "Calling WildBootTests.jl through fwildclusterboot"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Calling_WildBootTests_jl_through_fwildclusterboot}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
options(rmarkdown.html_vignette.check_title = FALSE)
```

```{r setup}
library(fwildclusterboot)
```

`fwildclusterboot` serves as an R port to the [WildBootTests.jl](https://github.com/droodman/WildBootTests.jl) implementation of the "fast" wild cluster bootstrap. 

## Installation and Set-Up

You can install Julia and `WildBootTests.jl` from within R via the `JuliaConnectoR.utils` package:

```{r, eval = FALSE, message = FALSE, warning = FALSE}
devtools::install_github("s3alfisc/JuliaConnectoR.utils")
library(JuliaConnectoR.utils)
install_julia() # install Julia
install_julia_packages("WildBootTests.jl") # install WildBootTests.jl
```

The `connect_julia_r()` function will help you to make sure that Julia and R find each other: 

```{r, eval = FALSE, message = FALSE, warning = FALSE}
connect_julia_r() # instructions to connect Julia and R
```


By default, the functions above install the most recent stable releases of both `Julia` and `WildBootTests.jl`.

If `Julia` and `WildBootTests.jl` are already installed, you can check if the installed versions match the `SystemRequirements` by running 

```{r, eval = FALSE, message = FALSE, warning = FALSE}
check_julia_system_requirements("fwildclusterboot")
```

Alternatively, you can install Julia from its official [https://julialang.org/downloads/](https://julialang.org/downloads/) and `WildBootTests.jl` via Julia's package manager. 

## The `boot_algo` function argument

You can run the wild cluster bootstrap through `WildBootTests.jl` via the `boot_algo` function argument: 

```{r, message = FALSE, warning = FALSE}
# load data set voters included in fwildclusterboot
data(voters)
# estimate the regression model via lm
lm_fit <- lm(proposition_vote ~ treatment + ideology1 + log_income + Q1_immigration , data = voters)
```

The first function call of `boottest()` will JIT-compile all required Julia code - this will take around 10-30 seconds. Note that if you re-start your R session, all Julia code will have to be re-compiled.

```{r, message = FALSE, warning = FALSE}
boot_lm <- boottest(lm_fit, clustid = "group_id1", param = "treatment", B = 9999, boot_algo = "WildBootTests.jl")
#summary(boot_lm)
```

If you decide that all your analyses should run through `WildBootTests.jl`, you can set a global variable via 

```{r, eval  = FALSE, , message = FALSE, warning = FALSE}
setBoottest_boot_algo("WildBootTests.jl")
```

All subsequent calls of `boottest()` will then automatically run `WildBootTests.jl` (unless you explicitly specify `boot_algo = "R"`).

```{r, eval = FALSE, message = FALSE, warning = FALSE}
# after setting setBoottest_boot_algo("WildBootTests.jl"), 
# the code below runs WildBootTests.jl without
# specifying 'boot_algo = "WildBootTests.jl"'
boottest(lm_fit, clustid = "group_id1", param = "treatment", B = 9999)
```


Note that if you want your analyses based on `WildBootTests.jl` to be reproducible, you need to set a seed for your `Julia` session. You can do so by calling 

```{r, eval = FALSE, message = FALSE, warning = FALSE}
set_julia_seed(87658)
```




