---
title: "Different Variants of the Wild Cluster Bootstrap"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

In a recent working paper, [MacKinnon, Nielsen and Webb (2022)](https://www.econ.queensu.ca/sites/econ.queensu.ca/files/wpaper/qed_wp_1485.pdf) suggest multiple new variants of the wild cluster bootstrap. The methods differ in a) how the bootstrap scores are computed and b) if a CRV1 or CRV3 variance matrix is used when computing bootstrapped t-statistics. 

All new bootstrap variants are implemented in `{fwildclusterboot}` and can be called by using the `boot_algo` function argument of `boottest()`. The implementation is still quite bare-bone: it only allows to test hypotheses of the form $\beta_k = 0$ vs $\beta_k \neq 0$, does not allow for regression weights or fixed effects, and further does not compute confidence intervals. 

To run e.g. the `WCR13` bootstrap, you would have to specify 

```{r setup}
library(fwildclusterboot)

N <- 1000
N_G1 <- 17
data <- fwildclusterboot:::create_data(
  N = N,
  N_G1 = N_G1,
  icc1 = 0.8,
  N_G2 = N_G1,
  icc2 = 0.8,
  numb_fe1 = 10,
  numb_fe2 = 5,
  seed = 41224,
  #seed = 123,
  weights = 1:N / N
)

lm_fit <- lm(
  proposition_vote ~ treatment + log_income, 
  data = data
)

wcr_algos <- c("R", "WCR11", "WCR13", "WCR31", "WCR33")
p_val <- t_stat <- 
  list()

for(x in wcr_algos){
  
  res <- 
    suppressWarnings(
        boottest(
      lm_fit, 
      param = ~treatment, 
      clustid = ~group_id1,
      B = 9999, 
      impose_null = TRUE,
      boot_algo = x, 
      seed = 123, 
      ssc = boot_ssc(
        adj = FALSE, 
        cluster.adj = FALSE
        )
      )
    )
  
  p_val[[x]] <- pval(res)
  t_stat[[x]] <- teststat(res)
  
}

df <- data.frame(
  "p-values" = unlist(p_val), 
  "t-statistics" = unlist(t_stat)
)

df

```
