---
title: "WildBootTests.jl"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(fwildclusterboot)
```


`fwildclusterboot` provides an R binding to the [WildBootTests.jl](https://github.com/droodman/WildBootTests.jl) implementation of the "fast" wild cluster bootstrap. 

## Installation and Set-Up

We recommend to install Julia from its official website [https://julialang.org/downloads/](https://julialang.org/downloads/).

The `JuliaConnectoR.utils` packages helps you to link R and Julia and to install `WildBootTests.jl` from within R. 
  
```{r, eval = FALSE, message = FALSE, warning = FALSE}
devtools::install_github("s3alfisc/JuliaConnectoR.utils")
library(JuliaConnectoR.utils)
connect_julia_r() # instructions to connect Julia and R
install_julia_packages("WildBootTests.jl") # install WildBootTests.jl
```

By default, the functions above install the most recent stable releases of `WildBootTests.jl`.

If `Julia` and `WildBootTests.jl` are already installed, you can check if the installed versions match the `SystemRequirements` by running 

```{r, eval = FALSE, message = FALSE, warning = FALSE}
check_julia_system_requirements("fwildclusterboot")
```

To set the number of threads for use in Julia, you can run 

```{r, warning = FALSE, message = FALSE, eval = FALSE}
set_julia_nthreads() # instructions to set nthreads for Julia
```

and follow the provided instructions. 

## The `boot_algo` function argument

You can run the wild cluster bootstrap through `WildBootTests.jl` via the `boot_algo` function argument: 

```{r, eval = FALSE, message = FALSE, warning = FALSE}
# load data set voters included in fwildclusterboot
data(voters)
# estimate the regression model via lm
lm_fit <- lm(proposition_vote ~ treatment + ideology1 + log_income + Q1_immigration , data = voters)
```

The first function call of `boottest()` will JIT-compile all required Julia code - this will take around 10-30 seconds. Note that if you re-start your R session, all Julia code will have to be re-compiled.

```{r, eval = FALSE, message = FALSE, warning = FALSE}
boot_lm <- boottest(lm_fit, clustid = "group_id1", param = "treatment", B = 9999, boot_algo = "WildBootTests.jl")
```

If you decide that all your analyses should run through `WildBootTests.jl`, you can set a global variable via 

```{r, eval  = FALSE, , message = FALSE, warning = FALSE}
setBoottest_boot_algo("WildBootTests.jl")
```

All subsequent calls of `boottest()` will then automatically run `WildBootTests.jl` (unless you explicitly specify `boot_algo = "R"`).

```{r, eval = FALSE, message = FALSE, warning = FALSE}
# after setting setBoottest_boot_algo("WildBootTests.jl"), 
# the code below runs WildBootTests.jl without
# specifying 'boot_algo = "WildBootTests.jl"'
boottest(lm_fit, clustid = "group_id1", param = "treatment", B = 9999)
```




